Precise predictions for same-sign W-bosons scattering at the LHC

### 2018-02-03 Draft/contributions.tex

At tree level, there are three contributions to the $W^+W^+$ production in association with two jets: the pure EW component $\mathcal{O}(\alpha^6)$, the interference $\mathcal{O}(\alphas\alpha^5)$ and the QCD background $\mathcal{O}(\alphas^2\alpha^4)$.
In the present section, the cross sections and distributions are obtained without applying the typical VBS cuts $m_{\Pj\Pj}$ and $|\Delta y_{\Pj\Pj}|$.
In Table~\ref{tab:LOscanXsec}, the cross section of the three contributions are displayed.
The EW, QCD, and interference contributions amount to $57\%$, $37\%$, and $6\%$ of the total inclusive cross section, respectively.
The QCD contribution does not posses external gluons due to charge conservation.
Thus the $\mathcal{O}(\alphas^2\alpha^4)$ diagrams only involve gluon exchange in the $t/u$-channel between the quark lines.
This results in a small contribution although the VBS cuts have not been imposed.
The interference between EW and QCD contributions is small, due to color suppression, but not negligible ($t/u$ interference with identical fermions).

@@

In Fig.~\ref{fig:mjjdyjj_1d} these three contributions are shown separately and summed in the differential distribution of the di-jet invariant mass $m_{\Pj\Pj}$ and the rapidity difference $|\Delta y_{\Pj\Pj}|$.
Interestingly, the EW contributions peaks around and invariant mass of about $80\GeV$.
These are due to contributions where the two jets originate from the decay of a W boson (see middle and right diagrams in Fig.~\ref{diag:LO}).
Note that these contributions are not present in calculation relying on the VBS approximation.
Also, the EW contribution becomes dominant for di-jet invariant mass larger than $500\GeV$ and for jet rapidity difference larger than $2.5$.

>>>>

At tree level, there are three contributions to the $W^+W^+$ production in association with two jets: the pure EW component $\mathcal{O}(\alpha^6)$, the interference $\mathcal{O}(\alphas\alpha^5)$ and the QCD background $\mathcal{O}(\alphas^2\alpha^4)$.
In the present section, the cross sections and distributions are obtained without applying the VBS cuts on $m_{\Pj\Pj}$ and $|\Delta y_{\Pj\Pj}|$.
In Tab.~\ref{tab:LOscanXsec}, the cross sections of the three contributions are displayed.
The EW, QCD, and interference contributions amount to $57\%$, $37\%$, and $6\%$ of the total inclusive cross section, respectively.
The QCD contribution does not posses external gluons due to charge conservation.
Thus the $\mathcal{O}(\alphas^2\alpha^4)$ diagrams only involve gluon exchange in the $t/u$-channel between the quark lines.
This results in a small contribution although the VBS cuts have not been imposed.
The interference between EW and QCD contributions is small, due to color suppression, but not negligible ($t/u$ interference with identical fermions).

@@

In Fig.~\ref{fig:mjjdyjj_1d} these three contributions are shown separately and summed in the differential distribution of the di-jet invariant mass $m_{\Pj\Pj}$ and the rapidity difference $|\Delta y_{\Pj\Pj}|$.
The EW contribution peaks around an invariant mass of about $80\GeV$.
These are due to diagrams where the two jets originate from the decay of a W boson (see middle and right diagrams in Fig.~\ref{diag:LO}).
Note that these contributions are not present in calculation relying on the VBS approximation.
Also, the EW contribution becomes dominant for di-jet invariant mass larger than $500\GeV$ and for jet rapidity difference larger than $2.5$.

### 2018-02-03 Draft/definition.tex

This process possesses three LO contributions of different orders.
The first one is of order $\mathcal{O}{\left(\alpha^{6}\right)}$ and is referred to as EW contributions.
In addition to typical VBS contributions as shown on the left of Fig.~\ref{diag:LO}, it also features $s$-channel contributions.
These types of contributions will play a particular role in the study of the various contributions in Section \ref{subsec:contributions}.
Some of them take the form of decay chains as for example the diagram represented in the middle of Fig.~\ref{diag:LO} while others are tri-boson contributions (right of Fig.~\ref{diag:LO}).
The VBS diagrams typically dominate the full process.
But all these contributions form a single gauge-invariant set of contributions and therefore cannot be separated.

The process can also be mediated via a gluon connecting the two quark lines while the W bosons are radiated from the quark lines.
These contributions are of order $\mathcal{O}{\left(\alphas^{2}\alpha^{4}\right)}$ and usually feature different kinematic behaviours than the EW contribution.
Nonetheless they share the same final state and therefore constitute an irreducible background.

Finally, due to the specific colour structure of these two classes of amplitudes, there exist non-zero interferences.
These are of order $\mathcal{O}{\left(\alpha_{\rm s}\alpha^{5}\right)}$ and are usually small but not negligible for realistic experimental set-ups \cite{Biedermann:2017bss}.

Usually, in experimental measurements, special VBS-cuts are designed in order to enhance the EW contribution over the QCD one.
These cuts are based on the fact that the two contributions have rather different kinematic.
The EW contribution is characterised by two jets with large rapidity in the peripheral region as well as a large invariant mass.
The two W bosons are mostly produced centrally.
This is in opposition with the QCD component which favours jets in the central region.
Therefore, the event selection usually involves rapidity-difference and invariance-mass cuts for the jets.
This will also be discussed in Section \ref{subsec:contributions}.

>>>>

This process possesses three LO contributions of different orders.
The first one is of order $\mathcal{O}{\left(\alpha^{6}\right)}$ and is referred to as EW contribution.
In addition to typical VBS contributions, as shown on the left of Fig.~\ref{diag:LO}, it also features $s$-channel contributions.
Note that we define $s$-, $t$-, and $u$-channel contributions by looking at the process which is contained by only looking at the quark lines.
The $s$-channels will play a particular role in the study of the various contributions in Sec.~\ref{subsec:contributions}.
Some of them take the form of decay chains, for example the diagram represented in the middle of Fig.~\ref{diag:LO}, while others are tri-boson contributions (right of Fig.~\ref{diag:LO}).
The VBS diagrams typically dominate the full process.
But all these contributions form a single gauge-invariant set of contributions and therefore cannot be separated.

The process can also be mediated via a gluon connecting the two quark lines while the W bosons are radiated off the quark lines.
These contributions are of order $\mathcal{O}{\left(\alphas^{2}\alpha^{4}\right)}$ and feature different kinematical behaviours than the EW contribution.
Nonetheless they share the same final state and therefore constitute an irreducible background.

Finally, due to the specific colour structure of these two classes of amplitudes, there exist non-zero interferences if only one quark family is involved.
These are of order $\mathcal{O}{\left(\alpha_{\rm s}\alpha^{5}\right)}$ and are usually small but not negligible for realistic experimental set-ups \cite{Biedermann:2017bss}.

In experimental measurements, special VBS-cuts are designed to enhance the EW contribution over the QCD one.
These cuts are based on the fact that the two contributions have different kinematical behaviour.
The EW contribution is characterised by two jets with large rapidities as well as a large invariant mass.
The two W bosons are mostly produced centrally.
This is in contrast to the QCD contribution which favours jets in the central region.
Therefore, the event selection usually involves rapidity-difference and invariance-mass cuts for the jets.
This will also be discussed in Sec.~\ref{subsec:contributions}.

### 2018-02-03 Draft/process.tex
References added

As mentioned previously, the EW contribution is dominated by the scattering of two W gauge bosons.
Therefore it is justified to approximate the full EW contributions simply by the VBS contributions.
Nonetheless, this set of contributions is not gauge invariant.
To make it gauge-invariant, one should project on-shell the incoming W boson.
Unfortunately, this momentum is space-like and thus a simple on-shell projection is not possible.
Instead, one can keep the W boson leg connected to the external quark line off-shell while the one connected to the scattering is put on-shell.
Then the polarisation of the gauge boson is accommodated following the implementations of Refs.~\cite{Kuss:1995yv,Accomando:2006hq}.
Such an approximation is usually called effective vector-boson approximation (EVBA).

A more refined approximation consists in considering all $t$- and $u$- diagrams and square them separately.
Nonetheless, different type of diagrams are not squared which amounts to neglecting interferences.
These interferences are expected to be small in the VBS fiducial region.
The $s$-channel contributions can still be left out.
This approximation is often called $t$-/$u$- approximation or even VBS approximation.
We will adopt the latter denomination in the following of the article.
Such an approximation is implemented at LO in the computer codes {\sc Bonsay} and {\sc Powheg}.
The physical origin of this approximation is that each proton emitting a quark represents two independent copies of the $SU_c\left(3\right)$ gauge group.

The squared matrix element of the $s$-channel contributions can then be added but all interferences between different kinematic channels are neglected.
This is what is done in {\sc VBFNLO}.

@@

Moving on to NLO accuracy, one can extend the approximation presented at LO.
The VBS approximation at NLO is simply extending the same approximation to real as well as to virtual corrections.
This is implemented in {\sc POWHEG}.
This approximation can be used in combination with a double-pole approximation [REF] for the virtual contribution.
This requires the computation of factorisable as well as non-factorisable corrections \cite{Dittmaier:2015bfe} separately.
Such an approximation is implemented in {\sc Bonsay}.
In {\sc VBFNLO}, the VBS approximation at NLO with $s$-channel contributions is implemented.

>>>>

As mentioned previously, the EW contribution is dominated by the scattering of two W gauge bosons.
Therefore it is justified to approximate the full EW contributions simply by the VBS contributions.
Nonetheless, this set of contributions is not gauge invariant.
To make it gauge-invariant, one should project on-shell the incoming W boson.
Unfortunately, this momentum is space-like and thus a simple on-shell projection is not possible.
Instead, one can keep the W boson leg connected to the external quark line off-shell while the one connected to the scattering is put on-shell.
Then the polarisation of the gauge boson is accommodated following the implementations of Refs.~\cite{Kuss:1995yv,Accomando:2006hq}.
Such an approximation is usually called effective vector-boson approximation (EVBA) \cite{Dawson:1984gx,Duncan:1985vj,Cahn:1983ip}.

A less crude approximation consists in considering all $t$- and $u$- diagrams and squaring them separately and neglecting their interferences.
These interferences are expected to be small in the VBS fiducial region.
The $s$-channel squared diagrams and interferences with $s$-channels are left out as well.
This approximation is often called $t$-/$u$- approximation, VBF, or even VBS approximation.
We will adopt the latter denomination in the following of the article.
Such an approximation is implemented at LO in the computer codes {\sc Bonsay} and {\sc Powheg}.
This approximation is gauge-invariant, which can be seen by considering the quark coming from different protons belonging to different $SU \left(3\right)$ gauge groups.

The squared matrix element of the $s$-channel contributions can added back, but all interferences between different kinematic channels are still neglected.
This is what is done in {\sc VBFNLO}.

@@

Moving on to NLO accuracy, one can extend the approximations presented at LO.
The VBS approximation at NLO is straightforward for the virtual contributions, for the real-contributions one must be careful about gluon-initiated processes\footnote{The initial gluon must not couple to the other initial quark, otherwise there are infrared divergences proportional to $s$-channels which do not match with the ones found in the virtual contributions.
The subset of diagrams where all couplings of the initial state gluon to initial state quark are neglected forms a gauge-invariant subset, with the same argument as presented above.
}.
This is implemented in {\sc POWHEG}.
This approximation can be used in combination with a double-pole approximation \cite{Dittmaier:2015bfe} for the virtual contribution.
Such an approximation is implemented in {\sc Bonsay}.
In {\sc VBFNLO}, the VBS approximation at NLO with $s$-channel contributions is implemented.
